---
title: "HW 1"
author: "Suwash Silwal"
output:
  pdf_document: default
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
xfun::session_info('rmarkdown')
```

## Problem 1: Let Y be a discrete random variable and has a binomial distribution with sample n = 20 and the sucessful probability $\theta$.

## 1.1 Probability mass function plot of Y, for $\theta$ = 0.05.
```{r}
n<- 20       # sample
p1<- 0.05    # theta
x<- 0:n      # Possible values of Y
bin1<- dbinom(x,n,p1)
plot(bin1,type="h", lwd=2, 
     main = "Binomial Distribution",
     col = "blue",
     ylab = "Pr(Y=x)",
     xlab = "(Y=x)")
```

## 1.2 Probability mass function plot of Y, for $\theta$ = 0.10.

```{r}
p2<- 0.10  # theta
bin2<- dbinom(x,n,p2)
plot(bin2,type="h", lwd=2, 
     main = "Binomial Distribution",
     col = "red",
     ylab = "Pr(Y=x)",
     xlab = "Y")
```

## 1.3 P(Y $\le$ 2), for $\theta$ = 0.05

```{r}
pbinom(2,n,p1)
```

## 1.4 P(Y $\le$ 2), for $\theta$ = 0.10

```{r}
pbinom(2,n,p2)
```

## Problem 2: Let $\theta$ be a continuous random variable and has a beta distribution with the parameter $\alpha$ = 2 and $\beta$ = 20.

## 2.1 Probability density function plot of $\theta$(x $\in$ (0,1))

```{r}
a<- 2      # alpha
b<- 20     # beta
x1<- seq(0,1,length = 100)   # random values between 0 and 1
beta1<- dbeta(x1,a,b)        
plot(x1,beta1, type="l", lwd=2, 
     main = "Beta Distribtuion",
     ylab = "Pr(Y=x)",
     col = "blue",
     xlab = "Y=x")
```

## 2.2 Cumulative function plot of $\theta$(x $\in$ (0,1))

```{r}
beta2<- pbeta(x1,a,b)  
plot(x1,beta2, type="l", lwd=2, 
     main = "CDF",
     ylab = "Pr(Y=x)",
     col = "red",
     xlab = "Y") 
```

## 2.3 Pr(0.05 $\le$ $\theta$ $\le$ 0.25)

```{r}
pbeta(0.25,a,b)- pbeta(0.05,a,b)
```

## 2.4 Pr($\theta$ $\le$ 0.75)

```{r}
pbeta(0.075,a,b)
```

## Problem 3: Let $\theta$ be a continuous random variable and has a beta distribution with the parameter $\alpha$ = 2 and $\beta$ = 40.

## 3.1 Probability density function plot of $\theta$(x $\in$ (0,1))

```{r}
a1<- 2     # alpha
b1<- 40    # beta
beta3<- dbeta(x1,a1,b1)     
plot(x1,beta3, type="l", lwd=2, 
     main = "Beta Distribtuion",
     ylab = "Pr(Y=x)",
     col = "blue",
     xlab = "Y")
```

## 3.2 Cumulative function plot of $\theta$(x $\in$ (0,1))

```{r}
beta4<- pbeta(x1,a1,b1)  
plot(x1,beta4, type="l", lwd=2, 
     main = "CDF",
     ylab = "Pr(Y=x)",
     col = "red",
     xlab = "Y") 
```

## 3.3 Pr(0.05 $\le$ $\theta$ $\le$ 0.25)

```{r}
pbeta(0.25,a1,b1)-pbeta(0.05,a1,b1)
```

## 2.4 Pr($\theta$ $\le$ 0.75)

```{r}
pbeta(0.075,a1,b1)
```

## Problem 4:

Given,

for $\theta = 1, Y \sim N(1,\sigma^2)$\
for $\theta = 2, Y \sim N(2,\sigma^2)$

And,

Pr($\theta$ = 1) = 0.25\
Pr($\theta$ = 2) = 0.75

\underline{1. Parameter space for $\theta$ and Sample space for Y}

\underline{Parameter space ($\Theta$):}

for $\theta$ = 1, $\Theta$ = {1, $\sigma$>0}\
for $\theta$ = 2, $\Theta$ = {2, $\sigma$>0}

\underline{Sample space (Y):}

for $\theta$ = 1, Y = y $\in$ (-$\infty$, $\infty$)\
for $\theta$ = 2, Y = y $\in$ (-$\infty$, $\infty$)

\underline{2. Prior distribution of $\theta$, p($\theta$)}

$\theta$ is discrete and the distribution of theta is:

p($\theta$) = 

\underline{3. Conditional distribution of Y/$\theta$, p(y/$\theta$)}

$Y/\theta \sim N(\mu_\theta, \sigma_\theta^2)$

\underline{4. The marginal distribution of Y, p(y)}

$p(y) = Pr(\theta=1) \times p(y/\theta = 1) + Pr(\theta=2) \times p(y/\theta = 2)$\
      $= 0.25 \times N(y|1, \sigma^2) + 0.75 \times N(y|2, \sigma^2)$

\underline{5. The marginal distribution of Y, p(y), when $\sigma$ = 2,}

$p(y) = 0.25 \times N(y |1,4) + 0.75 \times N(y|2,4)$

```{r}
sd<-2        # standard deviation 
theta1<- 1   # parameter 1
theta2<- 2   # parameter 2

y <- seq(-5,10,length = 1000) # random values between -5 to 10 

# marginal distribution of Y
mdf<- 0.25*dnorm(y,theta1,sd) + 0.75*dnorm(y,theta2,sd) 

# marginal distribution plot
plot(y,mdf,                  
     type="l", lwd=2, 
     main = "Marginal distribution of Y",
     ylab = "Pr(Y=y)",
     col = "blue",
     xlab = "Y=y")
```

\underline{6. The posterior probability Pr($\theta$=1|Y=1), when $\sigma$ = 2}

$Pr(\theta=1/y=1) = \frac{p(y=1|\theta =1)\times p(\theta=1)} {p(y=1)}$

```{r}
y1<- 1    # for r.v. Y = 1

# Probability of Y = 1 given theta = 1 
y1_theta<- 0.25*dnorm(y1,theta1,sd) 

# Marginal distribution for Y = 1 
mdf1<- 0.25*dnorm(y1,theta1,sd) + 0.75*dnorm(y1,theta2,sd) 

# Posterior probability Pr(theta=1|Y=1)
theta_y1<- y1_theta/mdf1  
theta_y1

```

\underline{7. The posterior distribution of $\theta, p(\theta|Y=1)$ for different $\sigma$}

```{r}
y_1 <- seq(-20,20,length = 1000)        # random values between -5 to 10 
sd_v<- c(0.5,2,4,6,15)                     # different values of s.d.
typ <- c("black","blue","red","green","violet")

# Marginal distribution
mdf_1<- 0.25*dnorm(y_1,theta1,sd_v[1]) + 0.75*dnorm(y_1,theta2,sd_v[1]) 

# Posterior probability Pr(theta|Y)
theta1_y <- (dnorm(y_1,theta1,sd_v[1])*0.25)/mdf_1   

# Posterior probability plot
plot(y_1,theta1_y,
     type="l", lwd=3, 
     main = expression(paste("Poesterior distribution of ", theta)),
     ylab = "Pr(Y=y)",
     col = typ[1],
     xlab = "Y=y")

for (i in 2:5){
  mdf_i<- 0.25*dnorm(y_1,theta1,sd_v[i]) + 0.75*dnorm(y_1,theta2,sd_v[i])
  thetai_y <- (dnorm(y_1,theta1,sd_v[i])*0.25)/mdf_i
  lines(y_1,thetai_y,col=typ[i])
}
legend("topright",col=typ,lty=1,
        c("0.5", "2", "4", "6","15"),
        title = expression(paste("s.d. " , (sigma))))
```

From the above plot, we can see that as $\sigma$ increases, the distribution tends to be uniform (because when $\sigma \rightarrow \infty$ the data contain no information)




